A Data-Driven Framework for Neural Field Modeling

### 2010-12-31 ltx/first_order_framework/response_to_reviews.tex

Parham: I made a copy for my edits so that we don't get conflicts.

\emph{This appendix provides a derivation for equations ?? and ??. The derivation is adapted from (Sanner et al., 1992, and Scerri et al., 2009). Consider the cortical sheet defined on the surface $\Omega$ with the sensors placed on a regular grid at positions $\mathbf{r}_n$ (where $n=0,...,n_y-1$ indexes the sensors) that are separated by the distance $\Delta_y$. The observations can be considered samples from a continuous field that is filtered (convolved) with the observation kernel described by
\begin{equation}
	y(\mathbf{r}_n)=\sum_{n=0}^{n_y-1}\delta(\mathbf r - \mathbf{r}_n ) v_m(\mathbf{r}),
\end{equation} 
where $v_m(\mathbf{r}) = v_t(\mathbf{r})\ast m(\mathbf{r})$. The Fourier transform of $y(\mathbf{r}_n)$ is
\begin{align}
 Y(\boldsymbol{\nu}) &= \frac{1}{\Delta_y^2}\sum_{n=0}^{n_y-1} \delta\left(\boldsymbol{\nu}-\frac{n}{\Delta_y}\right) \ast V_m(\boldsymbol \nu) \\
	&= \frac{1}{\Delta_y^2}V_m(\boldsymbol\nu)+\frac{1}{\Delta_y^2}\sum_{n=1}^{n_y-1} V_m\left(\boldsymbol{\nu}-\frac{n}{\Delta_y}\right).
\end{align} 
It should be clear that $Y(\boldsymbol{\nu})$ contains $V_m(\boldsymbol\nu)$ plus $n_y-2$ copies (aliases) of $V_m(\boldsymbol\nu)$ repeated on the regular lattice centered at $n\Delta_y^{-1}$. Due to the lowpass action of the observation kernel, the filtered dynamic field is band-limited with a cutoff frequency $\boldsymbol\nu_{cy}$. In order for $V_m(\boldsymbol\nu)$ to be disjointed from its aliases, the amount of shift $\Delta_y^{-1}$ must be at least $2\boldsymbol\nu_{cy}$, since each alias has a spectral range of
\begin{equation}
	\boldsymbol\nu = \left[\frac{n}{\Delta_y}-\boldsymbol{\nu}_{cy},\frac{n}{\Delta_y}+\boldsymbol{\nu}_{cy}\right].
\end{equation}
Therefore, to avoid aliasing the distance between observations must satisfy
\begin{equation}\label{eq:SamplingCriteria}
	\frac{1}{\Delta_y} \ge 2\nu_{cy}.
\end{equation}
Since the observation kernel is not an ideal filter an oversampling parameter should be included to further separate the aliases. Rearranging equation~\ref{eq:SamplingCriteria} and adding the oversampling parameter, $\rho$, gives
\begin{equation}\label{ap:Shannon}
	\Delta_y\le\frac{1}{2\rho\nu_{cy}},
\end{equation}
where $\rho \in \mathbb{R} \ge 1$. }

@@

    We thank the author for their encouraging review. Each point raised is dealt with in turn.

\begin{enumerate}
    \item The handling of the (to borrow a term from numerical weather forecasting) `sub-grid scale physics' is inadequate for nervous systems (see, e.g., equation 8 [The nonlinear IntegroDifference Equation]). I do very much appreciate the spectral estimation of spatial and temporal scales. But the model errors are lumped into an iid model error term that is not realistic. For instance, the flow of potassium in the extracellular spaces has a known dynamic that will affect network excitability, generating important sub-grid dynamics that is anything but iid. See, for instance, the example in Ullah et al PLoS Comp Biol 2010 where without such dynamics, state tracking and parameter estimation for a neural system fails. A beautiful overview of sub-grid effects in data assimilation is found in Kalnay's 2003 book. I would very much value the authors comments on these issues. I guess were I to respond to this question, that I might consider extending the model error term to an autoregressive model. But I will let the authors answer this question.\\

    \emph{talk about model order}

	\emph{Indeed there will always be a mismatch between the a mathematical model and the natural phenomena it is attempting to describe. The model proposed in our paper is modest in terms of the temporal dynamics, where the state transition is assumed to be first-order. However, the development of the framework for this simple model is a significant development that can be expanded to include longer range time dependencies. A major benefit in using a mean field model is the freedom to lump detailed parameters, such as specific ion concentrations, into other parameters such as the connectivity kernel coefficients. An extension to our framework that is currently in development is to allow for time-varying kernel parameters.  }

### 2011-01-01 ltx/first_order_framework/LaTeX_Manuscript.tex
added comments to address comment 1 from reviewer 2
added laplace method to extend the Greenie bastard
edits to paper and review letter

The firing rate of the presynaptic neurons is related to the post-synaptic membrane potential by the sigmoidal activation function 
\begin{equation}
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{1}{1 + \exp \left( \varsigma \left( v_0 - v\left(\mathbf{r}',t\right) \right) \right)}. 
\end{equation}
The parameter $v_0$ describes the firing threshold of the neural populations and $\varsigma$ governs the slope of the sigmoid. By substituting equation~\ref{RateBasedInteractions} into equation~\ref{SpikesToPotential} we get the spatiotemporal model 
\begin{equation}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) =
	\int_{-\infty}^t 
	h\left(t - t'\right) \int_\Omega
	w\left(\mathbf{r},\mathbf{r}'\right) 
	f\left( v\left( \mathbf{r}',t' \right)\right)
	\, d\mathbf{r}'dt'.
\end{equation}
Now we shall express the synaptic response kernel as a Green's function 
\begin{equation}
	\label{GreensFuncDef} \textrm{D}h\left( t \right) = \delta \left( t \right), 
\end{equation}
where $\textrm{D}=d/dt + \zeta$ is a temporal differential operator and $\delta(t)$ is the Dirac-delta function. This gives the standard form of the model
\begin{equation}
	\label{FinalFormContinuous} 
	\frac{dv\left( \mathbf{r},t \right)}{dt} + \zeta v\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)\, d\mathbf{r}'}. 
\end{equation}
To arrive at the integro-difference equation (IDE) form of the model, we discretize time using a first-order Euler method (see \ref{App:Time Discretization} for derivation) giving 
\begin{equation}
	\label{DiscreteTimeModel} 
	v_{t+T_s}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r}'\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\, d\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where $T_s$ is the time step, $\xi = 1-T_s\zeta $ and $e_t(\mathbf{r})$ is an $i.i.d.$ disturbance such that $e_t(\mathbf{r})\sim\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$. Here $\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$ denotes a zero mean Gaussian process with spatial covariance function $\gamma(\mathbf{r}-\mathbf{r}')$ (\cite{Rasmussen2005}). The disturbance is added to account for model uncertainty and unmodeled inputs. To simplify the notation, the index of the future time sample, $t+T_s$, shall be referred to as $t+1$ throughout the rest of the paper.

@@

This may hold:
\begin{equation}
	R_{\bar{y}_{t},\bar{y}_{t} } (\boldsymbol{\tau}) = \sigma_d \int_{\Omega}{ m(\mathbf{r}_n+\boldsymbol{\tau}-\mathbf{r}')\int_{\Omega}{ m(\mathbf{r}'_n-\mathbf{r}'') \gamma(\mathbf{r}'-\mathbf{r}'') d\mathbf{r}''} d\mathbf{r}'}.
\end{equation}
If it does then we can estimate the support of the disturbance using the correlation analysis if the bandwidth of the sensor is wider than the bandwidth of the disturbance.\\
the proof of the above\\
\begin{align}
	R_{y_{t},y_{t}}(\boldsymbol{\tau}) &= \mathbf{E}\left[y_{t}\left(\mathbf{r}_n+\boldsymbol{\tau}\right)y_{t}\left(\mathbf{r}_n\right)\right] \\
	&= \mathbf{E}\left[\left(\int_{\Omega}{ m\left(\mathbf{r}_n + \boldsymbol{\tau} - \mathbf{r}'\right) v_t\left(\mathbf{r}'\right)\, d\mathbf{r}'} + \boldsymbol{\varepsilon}_t\left(\mathbf{r}_n+\boldsymbol{\tau}\right)\right) \left(\int_{\Omega}{ m\left(\mathbf{r}_n - \mathbf{r}''\right) v_{t}\left(\mathbf{r}''\right) \, d\mathbf{r}''} + \boldsymbol{\varepsilon}_{t}\left(\mathbf{r}_n\right)\right) \right]\\
	&=\mathbf{E}\left[\int_{\Omega}{ m\left(\mathbf{r}_n + \boldsymbol{\tau} - \mathbf{r}'\right) v_t\left(\mathbf{r}'\right)\, d\mathbf{r}'} \int_{\Omega}{ m\left(\mathbf{r}_n - \mathbf{r}''\right) v_{t}\left(\mathbf{r}''\right) \, d\mathbf{r}''}\right]\\
	&=\iint_{\Omega}{ m\left(\mathbf{r}_n + \boldsymbol{\tau} - \mathbf{r}'\right) m\left(\mathbf{r}_n - \mathbf{r}''\right) \mathbf{E}\left[v_t\left(\mathbf{r}'\right) v_{t}\left(\mathbf{r}''\right)\right]\, d\mathbf{r}'}{ \, d\mathbf{r}''}\\
	&=\sigma_d\iint_{\Omega}{ m\left(\mathbf{r}_n + \boldsymbol{\tau} - \mathbf{r}'\right) m\left(\mathbf{r}_n - \mathbf{r}''\right)\gamma(\mathbf{r}'-\mathbf{r}'') \, d\mathbf{r}'}{ \, d\mathbf{r}''}\label{eq:ObsCorr}
\end{align}

@@

Let the spectral representation of the post-synaptic membrane voltage field at time $t$ be denoted by $V_t(\boldsymbol{\nu})$, where $\boldsymbol\nu$ is the spatial frequency. According to Shannon's sampling theorem, the field must be spatially band-limited for an accurate reconstruction using spatially discrete observations. However, an approximate reconstruction can be obtained if the field is approximately band-limited with 
\begin{equation}
	V_t(\boldsymbol{\nu}) \approx 0 ~ \ \forall \boldsymbol{\nu} > \boldsymbol{\nu}_c,
\end{equation}
where $\boldsymbol{\nu}_c$ is a cutoff frequency (typically taken as the -3~dB point). Given such a band-limited field, the distance between adjacent sensors, $\Delta_y$, must satisfy 
\begin{equation}
	\label{eq:MinimumSensorDistance} \Delta_y \leq \frac{1}{2\rho_y\boldsymbol{\nu}_{c}}, 
\end{equation}
where $\rho_y \in \mathbb{R} \ge 1$ is an oversampling parameter. This condition must be satisfied to avoid spatial spectral aliasing effects when reconstructing the hidden dynamic field, $v_t(\mathbf{r})$, using the sampled observations, $\mathbf{y}_t$.

@@

The frequency analysis specifies the minimum sensor arrangement that is required to capture the dominant cortical dynamics. As pointed out in numerous publications by Freeman and colleagues (see \citet{Freeman1987} for example), a detailed description of the spatial properties of responses to stimuli is essential in understanding cortical dynamics. Importantly, the spatial frequency analysis tools presented in this paper can be used to design electrophysiological recording systems to avoid spatial aliasing, ensuring the essential spatial dynamics can be observed. In addition, the frequency analysis highlights the relationship between the width of the sensors and the information in the data. This analysis is becoming increasingly relevant as recording systems are becoming more sophisticated with higher density electrodes \citep{Brinkmann2009}.

@@

@@

@@

\section{Observation Kernel}\label{App:ObsKernelExperiment}
The Gaussian shape of the observation kernel was verified experimentally. This was achieved by moving a dipole across a recording electrode submerged in saline. A photograph through a microscope of the submerged electrodes can be seen in Fig.~\ref{fig:SensorAndSource}. The measurement sensor was the most central electrode of a $5\times5$ Utah array (Blackrock Microsystems) which had electrode spacing of 400~$\mu$m. The current source was a twisted wire electrode (Plastics One Inc.) with tip-tip spacing of 60~$\mu$m and tip diameter of 200~$\mu$m. The twisted wire electrode was moved in the saline solution using a 3 dimensional manipulator (Sutter Instrument Company, MPC-100) in increments of 200~$\mu$m. The twisted wire electrode and the Utah array were co-registered visually using the microscope.
The stimulation waveform for the current source consisted of a 40~Hz sine wave with a duration of 1~s. The recording from the central electrode in the array was used to estimate the shape of the sensor kernel. To extract the amplitude of the signal, the data was first band-pass filtered with a center frequency of 40~Hz. The envelope of the signal was found by computing the mean instantaneous amplitude over the 1~s stimulation duration by
\begin{equation}
	a = \frac{1}{T}\sum_t{\sqrt{(y_t^2-\mathcal{H}y_t^2})}
\end{equation}
where $\mathcal{H}$ denotes the Hilbert transform operator, $t$ indexes samples and $T$ is the total number of samples in the time series. Fig.~\ref{fig:Ch3} shows the mean amplitude of the recorded envelope versus the distance between the twisted wire electrode and the recording electrode. A Gaussian fitted to the amplitudes is also shown. The figure shows that a Gaussian is indeed a good model for the sensor kernel in the homogeneous solution.
\section{Inner Product of Two \emph{n}-Dimensional Gaussians}\label{App:InnerProdOfGaussians}
The derivation in this section was used for calculating the matrix $\mathbf \Gamma$ from equation~\ref{eq:DefGamma}. Consider two Gaussian basis functions
\begin{equation}\label{eq:n_dimensional_Gaussian1}
 \varphi_i(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_i^2} (\mathbf r-\boldsymbol \mu_i)^\top(\mathbf r-\boldsymbol \mu_i})\right)
\end{equation}
and 
\begin{equation}\label{eq:n_dimensional_Gaussian2}
\varphi_j(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_j^2} (\mathbf r-\boldsymbol \mu_j)^\top(\mathbf r-\boldsymbol \mu_j})\right).
\end{equation}

The inner product of the two \emph{n}-dimensional Gaussians is defined as
\begin{equation}\label{eq:InofGaussianDefinition}
\varphi_i\otimes\varphi_j \triangleq \int_{\mathbb{R}^n}\varphi_i(\mathbf r)\varphi_j(\mathbf r)d\mathbf{r}.
A Gaussian basis function is symmetric with respect to its center, therefore
\begin{equation}\label{eq:GaussianSymmetry}
 \varphi_i(\boldsymbol \mu_i+\mathbf r)= \varphi_i(\boldsymbol \mu_i-\mathbf r).
Letting $\mathbf{r}= \mathbf{r}-\boldsymbol{\mu}_i$, equation~\ref{eq:GaussianSymmetry} becomes
\begin{equation}\label{eq:ConsequenceofGaussianSymmetry}
 \varphi_i(\mathbf r)= \varphi_i(2\boldsymbol \mu_i-\mathbf r).
Now substituting equation~\ref{eq:ConsequenceofGaussianSymmetry} into equation~\ref{eq:InofGaussianDefinition} gives 
\varphi_i\otimes\varphi_j&=\int_{\mathbb{R}^n}\varphi_i(\mathbf r)\varphi_j(\mathbf r)d\mathbf r \nonumber\\
&=\int_{\mathbb{R}^n}\varphi_i(2\boldsymbol \mu_i-\mathbf r)\varphi_j(\mathbf r)d\mathbf r \nonumber\\
&=(\varphi_i*\varphi_j)(2\boldsymbol \mu_i).
Now we solve the convolution using the formula derived in \ref{ap:ConvOfGaussians}, and evaluating it at $\mathbf r=2\boldsymbol\mu_i $ gives
 \varphi_i\otimes\varphi_j&=\left(\frac{\pi\sigma_i^2\sigma_j^2}{\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}} \nonumber \\
&\times\mathrm{exp}\left({-\frac{1}{\sigma_i^2+\sigma_j^2} (\boldsymbol\mu_i-\boldsymbol\mu_j)^\top(\boldsymbol\mu_i-\boldsymbol\mu_j)}\right).
This completes the derivation.
\section{Convolution of Two $n$-Dimensional Gaussians}\label{ap:ConvOfGaussians}

>>>>

The firing rate of the presynaptic neurons is related to the post-synaptic membrane potential by the sigmoidal activation function 
\begin{equation}
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{1}{1 + \exp \left( \varsigma \left( v_0 - v\left(\mathbf{r}',t\right) \right) \right)}. 
\end{equation}
The parameter $v_0$ describes the firing threshold of the neural populations and $\varsigma$ governs the slope of the sigmoid. By substituting equation~\ref{RateBasedInteractions} into equation~\ref{SpikesToPotential} we get the spatiotemporal model 
\begin{equation}
	\label{FullDoubleIntModel} v\left(\mathbf{r},t\right) =
	\int_{-\infty}^t 
	h\left(t - t'\right) \int_\Omega
	w\left(\mathbf{r},\mathbf{r}'\right) 
	f\left( v\left( \mathbf{r}',t' \right)\right)
	\, d\mathbf{r}'dt'.
\end{equation}

\dean{(reviewer 2, comment 3:) To arrive at the standard integro-differential equation form of the model, we use the fact that the synaptic response kernel is a Green's function of a linear differential equation defined by the differential operator $\textrm{D}=d/dt + \zeta$. A Green's function satisfies
\begin{equation}
	\label{GreensFuncDef} \textrm{D}h\left( t \right) = \delta \left( t \right), 
\end{equation} 
where $\delta(t)$ is the Dirac-delta function. To understand this property, one can take the Laplace transform of $\textrm{D}$ and $h(t)$, giving
\begin{align}
	\mathcal{L}\{h(t)\} &= \frac{1}{s+\zeta} \\
	\mathcal{L}\{\textrm{D}\} &= s+\zeta,
\end{align}
and then taking the inverse Laplace transform of $\mathcal{L}\{\textrm{D}\}$ applied to $\mathcal{L}\{h(t)\}$, giving
\begin{align}
	\mathcal{L}^{-1}\left\{\frac{s+\zeta}{s+\zeta}\right\} &= \mathcal{L}^{-1}\left\{1\right\} \\
	&= \delta \left( t \right).
\end{align}
Applying the differential operator $\textrm{D}$ to equation~\ref{FullDoubleIntModel} gives
\begin{align}
 \textrm{D}v\left(\mathbf r,t\right)&= \textrm{D}\left(h\ast g\right)\left(\mathbf r,t\right)\\
&=\left(\textrm{D}h\ast g\right)\left(\mathbf r,t\right)\\
&=\left(\delta \left(t\right)\ast g\right)\left(\mathbf r,t\right)\\
&=g\left(\mathbf r,t\right)
\end{align}
where $\ast$ denotes convolution operator.} This gives the standard form of the model
\begin{equation}
	\label{FinalFormContinuous} 
	\frac{dv\left( \mathbf{r},t \right)}{dt} + \zeta v\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)\, d\mathbf{r}'}. 
\end{equation}

To arrive at the integro-difference equation (IDE) form of the model, we discretize time using a first-order Euler method (see \ref{App:Time Discretization} for derivation) giving 
\begin{equation}
	\label{DiscreteTimeModel} 
	v_{t+T_s}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r}'\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\, d\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where $T_s$ is the time step, $\xi = 1-T_s\zeta $ and $e_t(\mathbf{r})$ is an $i.i.d.$ disturbance such that $e_t(\mathbf{r})\sim\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$. Here $\mathcal{GP}(\mathbf 0,\gamma(\mathbf{r}-\mathbf{r}'))$ denotes a zero mean Gaussian process with spatial covariance function $\gamma(\mathbf{r}-\mathbf{r}')$ (\cite{Rasmussen2005}). The disturbance is added to account for model uncertainty and unmodeled inputs. To simplify the notation, the index of the future time sample, $t+T_s$, shall be referred to as $t+1$ throughout the rest of the paper.

@@

This may hold:
\begin{equation}
	R_{\bar{y}_{t},\bar{y}_{t} } (\boldsymbol{\tau}) = \sigma_d \int_{\Omega}{ m(\mathbf{r}_n+\boldsymbol{\tau}-\mathbf{r}')\int_{\Omega}{ m(\mathbf{r}'_n-\mathbf{r}'') \gamma(\mathbf{r}'-\mathbf{r}'') d\mathbf{r}''} d\mathbf{r}'}.
\end{equation}
If it does then we can estimate the support of the disturbance using the correlation analysis if the bandwidth of the sensor is wider than the bandwidth of the disturbance.\\
the proof of the above\\
\begin{align}
	R_{y_{t},y_{t}}(\boldsymbol{\tau}) &= \mathbf{E}\left[y_{t}\left(\mathbf{r}_n+\boldsymbol{\tau}\right)y_{t}\left(\mathbf{r}_n\right)\right] \\
	&= \mathbf{E}\left[\left(\int_{\Omega}{ m\left(\mathbf{r}_n + \boldsymbol{\tau} - \mathbf{r}'\right) v_t\left(\mathbf{r}'\right)\, d\mathbf{r}'} + \boldsymbol{\varepsilon}_t\left(\mathbf{r}_n+\boldsymbol{\tau}\right)\right) \left(\int_{\Omega}{ m\left(\mathbf{r}_n - \mathbf{r}''\right) v_{t}\left(\mathbf{r}''\right) \, d\mathbf{r}''} + \boldsymbol{\varepsilon}_{t}\left(\mathbf{r}_n\right)\right) \right]\\
	&=\mathbf{E}\left[\int_{\Omega}{ m\left(\mathbf{r}_n + \boldsymbol{\tau} - \mathbf{r}'\right) v_t\left(\mathbf{r}'\right)\, d\mathbf{r}'} \int_{\Omega}{ m\left(\mathbf{r}_n - \mathbf{r}''\right) v_{t}\left(\mathbf{r}''\right) \, d\mathbf{r}''}\right] + \delta(\boldsymbol{\tau})\sigma_{\varepsilon}^2 \\
	&=\iint_{\Omega}{ m\left(\mathbf{r}_n + \boldsymbol{\tau} - \mathbf{r}'\right) m\left(\mathbf{r}_n - \mathbf{r}''\right) \mathbf{E}\left[v_t\left(\mathbf{r}'\right) v_{t}\left(\mathbf{r}''\right)\right]\, d\mathbf{r}'}{ \, d\mathbf{r}''} + \delta(\boldsymbol{\tau})\sigma_{\varepsilon}^2\\
	&=\sigma_d\iint_{\Omega}{ m\left(\mathbf{r}_n + \boldsymbol{\tau} - \mathbf{r}'\right) m\left(\mathbf{r}_n - \mathbf{r}''\right)\gamma(\mathbf{r}'-\mathbf{r}'') \, d\mathbf{r}'}{ \, d\mathbf{r}''} + \delta(\boldsymbol{\tau})\sigma_{\varepsilon}^2\label{eq:ObsCorr}
\end{align}

@@

Let the spectral representation of the post-synaptic membrane voltage field at time $t$ be denoted by $V_t(\boldsymbol{\nu})$, where $\boldsymbol\nu$ is the spatial frequency. According to Shannon's sampling theorem, the field must be spatially band-limited for an accurate reconstruction using spatially discrete observations. However, an approximate reconstruction can be obtained if the field is approximately band-limited with 
\begin{equation}
	V_t(\boldsymbol{\nu}) \approx 0 ~ \ \forall \boldsymbol{\nu} > \boldsymbol{\nu}_c,
\end{equation}
where $\boldsymbol{\nu}_c$ is a cutoff frequency (typically taken as the -3~dB point). Given such a band-limited field, the distance between adjacent sensors, $\Delta_y$, must satisfy 
\begin{equation}
	\label{eq:MinimumSensorDistance} \Delta_y \leq \frac{1}{2\rho_y\boldsymbol{\nu}_{c}}, 
\end{equation}
where $\rho_y \in \mathbb{R} \ge 1$ is an oversampling parameter. \dean{(reviewer 1, additional comment 3:) A derivation of the sampling theorem is provided in \ref{App:SamplingTheorem} that is based on the work of \citet{Sanner1992} and \citet{Scerri2009}}. This condition must be satisfied to avoid spatial spectral aliasing effects when reconstructing the hidden dynamic field, $v_t(\mathbf{r})$, using the sampled observations, $\mathbf{y}_t$.

@@

The frequency analysis specifies the minimum sensor arrangement that is required to capture the dominant cortical dynamics. As pointed out in numerous publications by Freeman and colleagues (see \citet{Freeman1987} for example), a detailed description of the spatial properties of responses to stimuli is essential in understanding cortical dynamics. Importantly, the spatial frequency analysis tools presented in this paper can be used to design \dean{(reviewer 1, item 3:) high density intracranial} electrophysiological recording systems to avoid spatial aliasing, ensuring the essential spatial dynamics can be observed. In addition, the frequency analysis highlights the relationship between the width of the sensors and the information in the data. This analysis is becoming increasingly relevant as recording systems are becoming more sophisticated with higher density electrodes \citep{Brinkmann2009}.

@@

\subsection{Real time implementation}
\dean{(comment 1 and 9 from reviewer 2:) The implementation of the framework has been described a being appropriate for an evoked-potential experimental protocol. Over the time duration of an experimental paradigm the connectivity and synaptic kernel parameters are assumed to be stationary. However, if the method is applied to a continuous recording over a longer time duration it is unlikely that the parameters will be stationary. If fact, it is likely that changes in the parameter values would be of most interest. 
To implement the algorithm in a real-time scenario the we would assume that the parameters are stationary over a time period (window) where the estimation is performed. The parameter of the previous window can be used to initialize parameters of the next window.  }

@@

\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig1.pdf}
\end{center}
\caption{}
\label{fig:Figure1}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig2.pdf}
\end{center}
\caption{}
\label{fig:Figure2}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig3.pdf}
\end{center}
\caption{}
\label{fig:Figure3}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig4.pdf}
\end{center}
\caption{}
\label{fig:Figure4}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig5.pdf}
\end{center}
\caption{}
\label{fig:Figure5}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig6.pdf}
\end{center}
\caption{}
\label{fig:Figure6}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig7.pdf}
\end{center}
\caption{}
\label{fig:Figure7}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig8.pdf}
\end{center}
\caption{}
\label{fig:Figure8}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig9.pdf}
\end{center}
\caption{}
\label{fig:Figure9}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig10.pdf}
\end{center}
\caption{}
\label{fig:Figure10}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig11.pdf}
\end{center}
\caption{}
\label{fig:Figure11}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig12.pdf}
\end{center}
\caption{}
\label{fig:Figure12}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig13.pdf}
\end{center}
\caption{}
\label{fig:SensorAndSource}
\end{figure*}
\clearpage
\newpage
\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/fig14.pdf}
\end{center}
\caption{}
\label{fig:Ch3}
\end{figure*}
\clearpage
\newpage

@@

\section{Observation Kernel}\label{App:ObsKernelExperiment}
The Gaussian shape of the observation kernel was verified experimentally. This was achieved by moving a dipole across a recording electrode submerged in saline. A photograph through a microscope of the submerged electrodes can be seen in Fig.~\ref{fig:SensorAndSource}. The measurement sensor was the most central electrode of a $5\times5$ Utah array (Blackrock Microsystems) which had electrode spacing of 400~$\mu$m. The current source was a twisted wire electrode (Plastics One Inc.) with tip-tip spacing of 60~$\mu$m and tip diameter of 200~$\mu$m. The twisted wire electrode was moved in the saline solution using a 3 dimensional manipulator (Sutter Instrument Company, MPC-100) in increments of 200~$\mu$m. The twisted wire electrode and the Utah array were co-registered visually using the microscope.
The stimulation waveform for the current source consisted of a 40~Hz sine wave with a duration of 1~s. The recording from the central electrode in the array was used to estimate the shape of the sensor kernel. To extract the amplitude of the signal, the data was first band-pass filtered with a center frequency of 40~Hz. The envelope of the signal was found by computing the mean instantaneous amplitude over the 1~s stimulation duration by
\begin{equation}
	a = \frac{1}{T}\sum_t{\sqrt{(y_t^2-\mathcal{H}y_t^2})}
\end{equation}
where $\mathcal{H}$ denotes the Hilbert transform operator, $t$ indexes samples and $T$ is the total number of samples in the time series. Fig.~\ref{fig:Ch3} shows the mean amplitude of the recorded envelope versus the distance between the twisted wire electrode and the recording electrode. A Gaussian fitted to the amplitudes is also shown. The figure shows that a Gaussian is indeed a good model for the sensor kernel in the homogeneous solution.

\section{Inner Product of Two \emph{n}-Dimensional Gaussians}\label{App:InnerProdOfGaussians}
The derivation in this section was used for calculating the matrix $\mathbf \Gamma$ from equation~\ref{eq:DefGamma}. Consider two Gaussian basis functions
\begin{equation}\label{eq:n_dimensional_Gaussian1}
 \varphi_i(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_i^2} (\mathbf r-\boldsymbol \mu_i)^\top(\mathbf r-\boldsymbol \mu_i})\right)
\end{equation}
and 
\begin{equation}\label{eq:n_dimensional_Gaussian2}
\varphi_j(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_j^2} (\mathbf r-\boldsymbol \mu_j)^\top(\mathbf r-\boldsymbol \mu_j})\right).
\end{equation}

The inner product of the two \emph{n}-dimensional Gaussians is defined as
\begin{equation}\label{eq:InofGaussianDefinition}
\varphi_i\otimes\varphi_j \triangleq \int_{\mathbb{R}^n}\varphi_i(\mathbf r)\varphi_j(\mathbf r)d\mathbf{r}.
\end{equation}
A Gaussian basis function is symmetric with respect to its center, therefore
\begin{equation}\label{eq:GaussianSymmetry}
 \varphi_i(\boldsymbol \mu_i+\mathbf r)= \varphi_i(\boldsymbol \mu_i-\mathbf r).
\end{equation}
Letting $\mathbf{r}= \mathbf{r}-\boldsymbol{\mu}_i$, equation~\ref{eq:GaussianSymmetry} becomes
\begin{equation}\label{eq:ConsequenceofGaussianSymmetry}
 \varphi_i(\mathbf r)= \varphi_i(2\boldsymbol \mu_i-\mathbf r).
\end{equation}
Now substituting equation~\ref{eq:ConsequenceofGaussianSymmetry} into equation~\ref{eq:InofGaussianDefinition} gives 
\begin{align}
\varphi_i\otimes\varphi_j&=\int_{\mathbb{R}^n}\varphi_i(\mathbf r)\varphi_j(\mathbf r)d\mathbf r \nonumber\\
&=\int_{\mathbb{R}^n}\varphi_i(2\boldsymbol \mu_i-\mathbf r)\varphi_j(\mathbf r)d\mathbf r \nonumber\\
&=(\varphi_i*\varphi_j)(2\boldsymbol \mu_i).
\end{align}
Now we solve the convolution using the formula derived in \ref{ap:ConvOfGaussians}, and evaluating it at $\mathbf r=2\boldsymbol\mu_i $ gives
\begin{align}
 \varphi_i\otimes\varphi_j&=\left(\frac{\pi\sigma_i^2\sigma_j^2}{\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}} \nonumber \\
&\times\mathrm{exp}\left({-\frac{1}{\sigma_i^2+\sigma_j^2} (\boldsymbol\mu_i-\boldsymbol\mu_j)^\top(\boldsymbol\mu_i-\boldsymbol\mu_j)}\right).
\end{align}
This completes the derivation.
